---
title: "Exploring quarto"
subtitle: "Baby steps into the quartoverse"
author: "Allison Horst"
date: "2022-04-09"
execute:
  warning: false
  message: false
---

## Getting started

First, let's create a new project and note some interesting differences (from .Rmd-verse).

1.  Restart RStudio
2.  New Project \> New Quarto Project

-   Note: the type of project that gets created
-   Note: the document extension
-   Note: Render button(instead of Knit)
-   Go ahead & render the .qmd!

## Add a new R code chunk

Now let's write some code in our .qmd & see what's different. If we're working in RStudio, the shortcut to add an R code chunk in a .qmd is the same (and default expected language is R).

-   Insert a new code chunk
-   Attach the `tidyverse` package

```{r}
library(tidyverse)
```

## Code chunk options - the hashpipe

We may want to hide some code chunk options. We can use the "hashpipe" `#|` to set code chunk options. For example, in the code chunk we just wrote, let's add the following:

-   `#| warning: false`
-   `#| message: false`

The other code chunk options we use (e.g. `echo`, `include`, `eval`) work similarly. Feel free to try them out!

## Global options

Update global options for code chunk execution in the front matter (up in the .qmd where the `title:`) is. That'll look something like this:

```{r}
#| eval: false
---
title: "Exploring quarto"
execution: 
  warning: false
  message: false
---
```

See more on execution options here: https://quarto.org/docs/computations/execution-options.html

## Outputs

So things seem pretty similar to .Rmd so far. Let's keep going with a little graph, then render.

```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg)) + 
  geom_point()
```

No surprises there! So what's new?

We can set options for outputs globally or within individual chunks (start figure labels with `fig-`...we'll see why in a bit.

```{r}
#| label: fig-faithful 
#| fig-cap: "Waiting time between Old Faithful eruptions."
#| fig-alt: Histogram showing the distribution of waiting times between Old Faithful eruptions. The distribution is bimodal, with a shorter duration peak around 55 minutes and a longer duration peak around 80 minutes. 

ggplot(data = faithful, aes(x = waiting)) +
  geom_histogram()
```

## But also you can just work in other languages:

```{python}
import seaborn as sns

df = sns.load_dataset("penguins")
sns.pairplot(df, hue = "species")
```

You can set code chunk execution options in Python (or other language) chunks the same way (using the hashpipe `#|`). 

## Some sugar

-   Check out the formatting when you add `author: "Name"` and `date: "XXXX-XX-XX"` to the front matter...pretty nice auto-formatting!
-   Cross-referencing (of figures and more!):

See @fig-faithful for a cool histogram.

- Learn more! [Quarto guide](https://quarto.org/docs/guide/)
